Stage-Splitting a Modal Language

### 2014-06-05 paper/paper.tex
more words about semantics.

We desire a semantics that meets the following goals,

@@

\begin{center}
\begin{tabular}{|c|p{2.5cm}|p{2.5cm}|} \hline
Judgment & Reads & Conditions \\ \hline 
$\isvalone e$ & ``$e$ is a $\bbone$-value under context $\Xi$'' & ... \\ \hline 
$\Xi \vdash \diaonesub$ & ``$e$ evaluates to future table $\gamma$ and value $v$'' 
& $\typesone e A$ 
	\newline $\isvalone v$ 
	\newline ... \\ \hline 
$\diatwosub$ & ``$e$ speculates to $q$'' & ... \\ \hline 
\end{tabular}
\end{center}

@@

We meet our first and second goals (show examples).

>>>>

We desire a semantics that meets the following goals:

@@

\begin{center}
\begin{tabular}{|c|p{2.5cm}|p{2.5cm}|} \hline
Judgment & Reads & Conditions \\ \hline 
$\isvalone e$ & ``$e$ is a $\bbone$-value under context $\Xi$'' & ... \\ \hline 
$\Xi \vdash \diaonesub$ & ``$e$ evaluates to future table $\gamma$ and value $v$'' 
& $\typesone e A$ 
	\newline $\isvalone v$ 
	\newline ... \\ \hline 
$\diatwosub$ & ``$e$ speculates to $q$'' & ... \\ \hline 
...&...&... \\ \hline
\end{tabular}
\end{center}

Evaluation and speculation are the main judgments here, the rest being largely administrative.  The evaluation judgment operates on stage-1 code whereas the speculation operates on stage-2 code.  Since $\next$ and $\prev$ are the crossover points between 1-code and 2-code, they are correspondingly the only places where the evaluation and speculation judgments depend on the other.

We can think of our evaluation as proceeding in the standard way for stage-1 code. But when the evaluator encounters stage-2 code, it places that code off to the side in a table, and then keeps a reference to the table entry.  These references are then passed around as stage-1 values, and the tables are managed accordingly.

The evaluation judgment is very similar to standard call-by-value evaluation, as goal 1 would suggest.  The input to evaluation is a stage-1 expression (usually $e$), as well as two administrative contexts ($\Gamma$ and $\Xi$), covered later.  Evaluation has two outputs: the {\em future table} (usually $\xi$) and the {\em partial value} (usually $v$).  We cover the latter first.  The partial value is essentially the result of the first-stage portion of $e$, and must be a 1-value of the same type as $e$.  As usual, this means that $v$ must be composed only of base primitives, tuples, injections, and lambdas (corresponding to base types, products, sums, and functions).  Analogously, the value corresponding to $\fut$ is a construct called a {\em hatted variable} (written $\hat y$), which signifies a reference to some stage-2 computation.  Those stage-2 computations are held in

@@

We meet our first (discuss) and second goals (show examples).
